% =====================================================================
% AAAI-27 SUBMISSION SKELETON  (AI Alignment special track)
% Retarget of "Objective Decoupling in Social RL".
%
% HOW TO USE:
%  - This compiles standalone (article). For submission, port the body into
%    the AAAI-27 Author Kit style (aaai27.sty). Do NOT submit in article class.
%  - Hard limit: 7 pages of technical content + unlimited references.
%    Full proofs and extra figures go in the separate technical appendix
%    (reuse ESA_corrections_and_upgrades.tex as the appendix source).
%  - Every \textcolor{red}{[RUN]} is an experiment YOU must run. The only
%    real numbers below are the two synthetic checks already produced; they
%    are marked [illustrative]. Everything else is a placeholder.
%  - \citep keys assume your existing .bib plus the entries added in the
%    corrections file. Standalone compile will warn on undefined cites.
% =====================================================================
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,bm}
\usepackage{enumitem,natbib,xcolor,hyperref,booktabs}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}\newtheorem{definition}{Definition}
\theoremstyle{definition}\newtheorem{assumption}{Assumption}

\newcommand{\Rstar}{R^{\ast}}\newcommand{\astar}{a^{\ast}}
\newcommand{\asoc}{a_{\mathrm{soc}}}\newcommand{\A}{\mathcal{A}}
\newcommand{\Cov}{\mathcal{C}}\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathcal{V}}\newcommand{\N}{\mathcal{N}}
\newcommand{\run}[1]{\textcolor{red}{[RUN: #1]}}
\newcommand{\note}[1]{\textcolor{blue}{[#1]}}

\title{\bf The Coverage Boundary of Source Auditing:\\ When Verifiable Rewards Recover Ground Truth from Biased Feedback}
\author{Majid Ghasemi \and Mark Crowley\\ University of Waterloo}
\date{}

\begin{document}
\maketitle
% Alternative titles to consider:
%  - "When Can Feedback Audits Recover Ground Truth? A Coverage Characterization"
%  - "Auditing the Judges: Coverage-Bounded Recovery under Systematic Evaluator Bias"

\begin{abstract}
\noindent
Alignment from human feedback assumes the feedback \emph{source} is a faithful channel to the
objective. We call this the Trusted-Feedback Assumption and show it fails under systematic evaluator
bias, producing \emph{objective decoupling}: the learned objective separates from ground truth with
probability one, and gold-standard consensus (e.g.\ Dawid--Skene) collapses once the bias is
majority-held. We ask when an agent can recover by auditing sources against a sparse reference. Our
main result is a coverage characterization: recovery is possible \emph{if and only if} the reference
covers the region where bias could flip the decision. Outside that region no source-auditing method
succeeds, and no statistic of the reports alone helps. We give a matching algorithm, \textbf{ESA},
that audits \emph{actively} (a horizon-independent number of queries), tracks trust per region, and
fails safe to abstention under total corruption. Instantiating the reference with verifiable
subtasks (RLVR) supplies the audit for free where a ground truth exists and reduces recovery on
non-verifiable tasks to a \emph{measurable} cross-domain reliability assumption. On a real LLM grader
panel with a GSM8K verifier, ESA recovers ground truth on non-verifiable TruthfulQA at $80\%$ where
every majority-trusting aggregator collapses to $\le 1\%$; the same experiments delimit the method
sharply --- an adversary strategic about the (public) audit boundary is maximally trusted rather than
caught, and the mechanism requires \emph{cardinal} feedback, inverting under ordinal elicitation.
\end{abstract}

% =====================================================================
\section{Introduction}
% ~1 page. Lead with the assumption, the failure, and the coverage thesis.
% Reframed via verifiable rewards from the first paragraph (fixes R4-w1/w4).
Reinforcement learning from human feedback treats annotations as a noisy but faithful view of a
latent objective $\Rstar$. This rests on an implicit commitment we name the \textbf{Trusted-Feedback
Assumption}: the feedback \emph{source} reports $\Rstar$ up to zero-mean noise. It is orthogonal to
the three dogmas of \citet{abel2024three}, which concern agent modeling, adaptation, and the reward
hypothesis, not source reliability.

In social settings the assumption fails. Feedback is produced by evaluators who may carry systematic
bias: sycophancy, effort-minimizing shortcuts, or coordinated manipulation. We show that under naive
aggregation this yields \emph{objective decoupling} (Def.~\ref{def:decoupling}): the learned
objective separates from $\Rstar$, and no-regret learning on the biased signal incurs linear latent
regret. Consensus truth discovery does not save it; once bias is majority-held, methods that assume
a mostly-honest crowd converge to the majority's decoupled objective.

The natural fix is to audit sources against a sparse, trusted reference rather than trust the crowd.
The central question of this paper is when that works. Our answer is sharp and includes a hard limit
on the method itself:

\smallskip
\noindent\emph{Recovery by source auditing is possible if and only if the reference covers the
region where bias could flip the decision.}
\smallskip

We prove both directions (Thms.~\ref{thm:lower},~\ref{thm:upper}) and show the limit is not specific
to our method: no function of the reports alone separates a truthful majority from a coherent
colluding one (Cor.~\ref{cor:internal}). This reframes the contribution away from ``beating
consensus'' and toward characterizing the boundary of feedback auditing.

We then make the reference concrete and cheap. Where correctness is computable, a verifier already
exists (RLVR; \citealp{lambert2024tulu,guo2025deepseekr1}): math answer-matching, unit tests,
checkable constraints. On that verifiable region the audit is dense and free, and recovery on the
non-verifiable region reduces to a \emph{measurable} cross-domain reliability assumption
(Prop.~\ref{prop:rlvr}). This scopes ``ground truth'' to where it genuinely exists and removes the
need to invoke a contested ``true human intent.''

\paragraph{Contributions.}
\begin{enumerate}[nosep,leftmargin=1.4em]
\item The Trusted-Feedback Assumption and an exact, margin-form condition for objective decoupling
(\S\ref{sec:theory}).
\item A coverage characterization of recovery: matching necessity (Thm.~\ref{thm:lower}, via a
two-point construction) and sufficiency (Thm.~\ref{thm:upper}), plus an internal-only impossibility
corollary (Cor.~\ref{cor:internal}).
\item \textbf{ESA}: active auditing with a horizon-independent budget (Thm.~\ref{thm:audit}),
per-region trust, an absolute-distrust fail-safe, and correlation-based audit routing
(\S\ref{sec:algo}).
\item A verifiable-subtask instantiation (RLVR) that supplies the reference for free and turns the
key assumption into a measured quantity (\S\ref{sec:rlvr}).
\item Validation on synthetic bandits and a real LLM grader panel: recovery under coverage, the
predicted failure boundary, and two sharp scope conditions (a strategic audit-boundary adversary that
is maximally trusted, and a cardinal-elicitation requirement) (\S\ref{sec:exp}).
\end{enumerate}

% =====================================================================
\section{Related Work}\label{sec:related}
% ~0.5 page. Condensed. All the fixes from the corrections file, one pass.
\paragraph{Robust reward modeling.} RRM \citep{liu2024rrm} disentangles reward from spurious
artifacts; distributionally robust DPO (WDPO, KLDPO) \citep{xu2025robust} guards the objective
against annotation shift. Both assume sparse corruption and a trustworthy bulk; we treat
\emph{majority-held, systematic} bias, where no clean bulk exists. They are also \emph{preference}
(ordinal) methods, and our modality result (\S\ref{sec:exp}) shows ordinal elicitation actively hides
the dispositional bias we study --- a lenient, correctly-ranking sycophant has \emph{lower} ordinal
loss than a truthful evaluator. We therefore compare against scalar-reward analogues of these methods
on our cardinal panel (they still fail under a systematic majority) rather than the original
preference-optimization objectives, which target a different corruption model on a different modality.

\paragraph{Truth discovery and gold questions.} Weighting annotators by verified ``gold'' items is
the standard way consensus survives a wrong majority \citep{dawid1979,li2016survey}. Our
contribution is not the anchor idea. It is the online, non-stationary RL loop, the partial-coverage
regime, and the resulting coverage characterization.

\paragraph{Robust statistics.} Overcoming an adversarial majority with a small trusted set is the
semi-verified / list-decodable problem \citep{charikar2017untrusted,steinhardt2018resilience};
reading the reference as the verified set gives the cleanest account of what it buys.

\paragraph{Spectral crowdsourcing.} Oracle-less bloc discovery draws on
\citep{karger2011iterative,zhang2014spectral,ghosh2011moderators}; we use it for partitioning only.

\paragraph{Verifiable rewards (RLVR).} Where correctness is computable, the reward is a deterministic
verifier \citep{lambert2024tulu,guo2025deepseekr1,shao2024deepseekmath}; DeepSeek-R1 chose rule-based
over neural rewards to avoid reward hacking. Non-cooperative crowdworker behavior is documented in
\citep{downs2010gaming,kazai2011worker}. \note{cite Su et al. 2025 for cross-domain verifier
agreement.}

% =====================================================================
\section{Setup}\label{sec:setup}
% ~1 page. Bandit for theory (clean, provable); MDP is empirical only.
We formalize the theory in the (contextual) bandit / single-state Social MDP; the MDP experiments
(\S\ref{sec:exp}) are empirical. Actions $a\in\A$, $|\A|=K$. Evaluator $m$ reports
\[
y_m(a) = \Rstar(a) + b_m(a) + \epsilon,\qquad \epsilon\ \sigma\text{-subgaussian},
\]
with $b_m$ the systematic bias (truthful $b\!\approx\!0$; sycophantic $b\propto$ prior; adversarial
$b\propto -\Rstar$). We stress that $y_m$ is the output of an evaluator \emph{policy}, not a noisy
sensor reading of $\Rstar$: the Social MDP models intent, so bias is directional, not symmetric.

\begin{definition}[Reference oracle with coverage]\label{def:reference}
A reference is informative only on a coverage set $\Cov\subseteq\A$. When queried (prob.
$p_{\mathrm{ref}}$) at $a\in\Cov$ it returns $z(a)=\Rstar(a)+\xi$, $\xi\ \sigma_z$-subgaussian; off
$\Cov$ it is uninformative. Typically $\Cov$ is a strict subset: a safety constraint certifies
``not lava'' but does not rank ``hop'' vs ``stand.''
\end{definition}

\begin{definition}[Objective decoupling]\label{def:decoupling}
With $J(\pi)$ the latent return, decoupling holds if $\Delta=J(\pi^\ast)-\E[J(\hat\pi)]>0$ as
$T\to\infty$, where $\hat\pi$ is learned from the aggregated feedback.
\end{definition}

ESA maintains per-region trust $w_m^{(g)}$ (Def.~\ref{def:regiontrust}) and aggregates
$\hat r(a)=\sum_m w_m y_m(a)$. Dominance is derived, not assumed: on $\Cov$, the expected
reference-loss gap between any biased and any truthful evaluator is
$\gamma_\Cov \ge \min_{m\in\text{bias}}\E_\Cov|b_m| - 2\E|\epsilon-\xi|$, so it holds exactly when
biases exceed the noise floor on the covered region (App.~\ref{app:dominance}).

% =====================================================================
\section{The Coverage Boundary}\label{sec:theory}
% ~2 pages. The core. Statements here; full proofs in appendix.
\begin{proposition}[Decoupling condition]\label{prop:decoupling}
With naive-mean bias $\bar B(a)=\frac1M\sum_m b_m(a)$ and latent gap $\Delta(a)=\Rstar(\astar)-\Rstar(a)$,
decoupling ($\asoc\neq\astar$) holds iff $\max_{a\neq\astar}[\bar B(a)-\bar B(\astar)-\Delta(a)]>0$.
Then any no-regret learner on the naive-mean signal incurs latent regret $\Omega(T)$.
\end{proposition}

\begin{lemma}[Trust error to policy error]\label{lem:bridge}
$|\hat r(a)-\Rstar(a)|\le \beta_\ast + W_{\mathrm{bias}}(t)B_{\max} + O_P(\sigma/\sqrt{N(a)})$,
with $\beta_\ast$ the residual truthful bias. If this is below $\Delta_{\min}/2$ for all $a$, the
empirical argmax is $\astar$ and the base learner converges to it.
\end{lemma}

\begin{theorem}[Coverage is necessary]\label{thm:lower}
Fix an uncovered arm $a_0\notin\Cov$. Two instances: $P_0$ all-truthful with $\Rstar(a_0)$ high
($a_0=\astar$); $P_1$ all evaluators biased by $\Delta$ on $a_0$ with $\Rstar(a_0)$ low
($a_0\neq\astar$). The two induce an identical observation law, so for every algorithm
$\max\{\E_{P_0}[R^{\mathrm{lat}}_T],\E_{P_1}[R^{\mathrm{lat}}_T]\}\ge\frac{\Delta}{2}T$.
\end{theorem}

\begin{corollary}[No internal-only rule survives a coherent majority]\label{cor:internal}
Any rule that is a function of the report matrix $Y$ alone (peer agreement, residual correlation,
spectral) inherits the same $\Omega(T)$ bound: reports cannot separate a truthful majority from a
colluding one.
\end{corollary}

\begin{assumption}[Coverage margin]\label{ass:coverage}
$\Cov$ contains every action whose aggregated observed value can exceed $\Rstar(\astar)$, and
dominance holds on $\Cov$ with margin $\gamma>0$.
\end{assumption}

\begin{theorem}[Coverage is sufficient]\label{thm:upper}
Under Assumption~\ref{ass:coverage}, ESA drives $W_{\mathrm{bias}}\!\to\!0$; after a burn-in of
$O\!\big((\eta\gamma p_{\mathrm{ref}}\rho)^{-1}\log(M/\delta)\big)$ reference-effective rounds the
aggregation error falls below $\Delta_{\min}/2$ and the base learner attains $\tilde O(\sqrt T)$
latent regret.
\end{theorem}

\noindent\emph{Together, Thms.~\ref{thm:lower} and \ref{thm:upper} match on the coverage axis:
coverage of the divergence region is necessary and sufficient.} This is the paper's characterization;
``beats Dawid--Skene'' is a corollary, not the point.

\begin{proposition}[Fail-safe under total corruption]\label{prop:failsafe}
Add an absolute rule: if every evaluator's reference loss exceeds a threshold $\tau$ (with
$\E|\epsilon-\xi|<\tau$), distrust all and act by $\pi_{\mathrm{safe}}$ (follow $z$ where certified,
abstain elsewhere). Then under homogeneous bias $|b|>\tau+\E|\epsilon-\xi|$ (e.g.\ 100\% bias) ESA
does not converge to $\asoc$; it fails toward abstention, not toward the lie.
\end{proposition}

\begin{proposition}[Strategic adversary; Stackelberg]\label{prop:stackelberg}
Against the committed trust rule, any bias retaining nonvanishing influence satisfies
$|\delta|\le\min(2\E|\epsilon-\xi|,\,\tau+\E|\epsilon-\xi|)=O(\sigma_{\mathrm{eff}})$: sustained bias
cannot exceed the noise floor.
\end{proposition}

Full proofs: App.~\ref{app:proofs}.

% =====================================================================
\section{Algorithm: ESA}\label{sec:algo}
% ~0.5 page + algorithm float (can go to appendix if space is tight).
\begin{definition}[Per-region trust]\label{def:regiontrust}
Partition the auditable space into regions $g$ (one per competitive action minimum) and keep
$w_m^{(g)}$, updated only by audits in $g$. This prevents an evaluator's bias in one region from
decaying its weight in another.
\end{definition}

ESA (i) selects actions by UCB/policy on $\hat r$, (ii) audits \emph{actively} (below), (iii)
updates per-region trust by multiplicative weights on $\ell_m=|y_m-z|$, (iv) applies the
absolute-distrust rule, and (v) passes $\hat r$ to the base optimizer (Q-learning or PPO). Full
pseudocode: App.~\ref{app:algo}.

\paragraph{Active auditing.} Audit only where a region is unresolved (trust not yet concentrated) and
a competitive action remains; stop once resolved.
\begin{theorem}[Audit complexity]\label{thm:audit}
Active auditing recovers $\astar$ with $O(K_{\mathrm{eff}}\,\gamma^{-2}\log(K_{\mathrm{eff}}/\delta))$
reference queries, \emph{independent of $T$}, where $K_{\mathrm{eff}}$ is the number of
ever-competitive actions. The fixed-rate rule uses $\Theta(p_{\mathrm{ref}}T)$.
\end{theorem}

\paragraph{Correlation routing.} Report-covariance clustering recovers the evaluator partition into
blocs with no oracle; when bias is bloc-coherent this cuts $K_{\mathrm{eff}}$ to the bloc count $r$
(App.~\ref{app:proofs}). It supplies the partition only; labeling still needs the audit
(Cor.~\ref{cor:internal}). A covariance shift re-flags a resolved region, handling strategic
adaptation and contagion.

% =====================================================================
\section{Verifiable-Subtask Anchoring}\label{sec:rlvr}
% ~0.5 page. The framing fix + free reference.
Replace hand-designed axioms with the verifier that already exists wherever correctness is
computable. Partition into a verifiable region $\V$ (checker present; $\sigma_z\!\approx\!0$,
effective $p_{\mathrm{ref}}=1$) and a non-verifiable region $\N$. On $\V$ the sparse-audit limitation
vanishes: $\V$ is the coverage set, dense and free.

\begin{proposition}[Verifiable-to-non-verifiable transfer]\label{prop:rlvr}
If cross-region reliability drift is bounded, $|b_m^{\N}-b_m^{\V}|\le\Delta_{\mathrm{drift}}$, then
trust learned on $\V$ yields aggregation error on $\N$ at most
$\beta_\ast^{\V}+\Delta_{\mathrm{drift}}+W_{\mathrm{bias}}B_{\max}$, and recovery on $\N$ holds when
this is below $\Delta_{\min}/2$. If drift is unbounded, Thm.~\ref{thm:lower} with $\Cov=\V$ applies:
no verifiable-subtask method recovers the $\N$-optimum.
\end{proposition}

$\Delta_{\mathrm{drift}}$ is measurable: hold out verifiable prompts that also carry human labels and
test whether reliability on the verifiable part predicts reliability elsewhere. This turns the key
assumption into an experiment (\S\ref{sec:exp}), and it scopes recovery to $\V$, where a ground truth
genuinely exists, rather than to a contested global intent. Verifiers are imperfect (gameable), so
we keep $\sigma_z>0$ on $\V$.

% =====================================================================
\section{Experiments}\label{sec:exp}
% ~1.5 pages. HONEST: only the two synthetic checks are real. Rest is [RUN].
\paragraph{Q1: Recovery under coverage (MDP).} Gridworld sycophant-trap (10$\times$10, 80\% lazy
sycophants, 15 seeds): standard RL and median-robust aggregation are fully trapped
($P(\text{proxy goal})=1.00$), and \emph{global}-trust ESA is nearly so ($0.93$) --- because the
sycophants are honest in most cells, so the rare divergence is diluted under one weight vector; only
\emph{per-region} ESA escapes ($0.07$). Hopper-v4 under an 80\% velocity-penalizing majority (5 seeds,
PPO): per-region ESA attains latent return $2.37\pm0.27$ vs.\ standard PPO $0.97\pm0.00$ and an
IRL/GAIL imitation baseline $0.98$; a scale-free-network contagion variant shows ESA quarantining the
highest-degree influencer (trust $\to 0.005$) at 95\% infection.

\paragraph{Q2: The predicted failure boundary (load-bearing).} Bandit with the reference covering only
part of the space. When the divergence arm is \emph{covered}, ESA recovers the optimum $100\%$ of
seeds; when it is \emph{uncovered} (sycophants lie on an unaudited optimum), ESA recovers $0\%$, as
does the mean --- matching Thm.~\ref{thm:lower}. At $100\%$ homogeneous bias, pure trust-auditing also
fails ($0\%$); only the absolute-distrust fail-safe recovers, and only under coverage.

\paragraph{Q3: Audit economy.} In a $K{=}10$ bandit with an 80\% sycophantic majority, fixed-rate
($p_{\mathrm{ref}}=0.1$) auditing grows as $\Theta(p_{\mathrm{ref}}T)$ ($\approx 8000$ audits at
$T=8{\cdot}10^4$) while active auditing stays flat at $\approx 32$ across $T\in[10^3,8{\cdot}10^4]$;
both recover $\astar$ --- the horizon-independence of Thm.~\ref{thm:audit}.

\paragraph{Q4: Bloc partition.} On the \emph{synthetic} bandit, oracle-less covariance clustering
recovers a colluding-majority partition at $0.7$--$1.0$ accuracy (and cannot label which bloc is
honest, per Cor.~\ref{cor:internal}); a drift alarm fires within 25 steps of a bloc forming.
\emph{Negative transfer to real graders:} on the LLM panel the partition does \emph{not} separate
(accuracy $0.70$, no clean split) because a one-model-per-type panel has no within-bloc diversity to
detect; we report this and keep Q4 synthetic-only.

\paragraph{Q5: RLVR instantiation and the transfer boundary (real graders).} Verifier $z=$ GSM8K
exact-match on $\V$; $\N=$ TruthfulQA; panel of 2 honest (Qwen2.5-32B) + 8 sycophant (Qwen2.5-7B)
graders; Qwen2.5-32B as an independent gold on $\N$; $N=100$, $95\%$ bootstrap CIs.
Trust learned on $\V$ transfers to $\N$: ESA recovers $80\%\,[72,87]$ where the sycophantic majority
captures the naive mean at $1\%$. Three findings sharpen the picture. \emph{(a) Coupling is
structural:} a submitted answer is itself an expressed belief, so a dispositional sycophant is
detectable even on $\V$ --- via grading \emph{leniency} (it scores a wrong math answer $6.3/10$ vs.\
honest $1.9$), which the exact verifier punishes. \emph{(b) The strategic adversary:} a sycophant
\emph{instructed to grade math strictly} (wrong$\to 0.08$) is honest-on-$\V$/biased-on-$\N$ --- the
Thm.~\ref{thm:lower} adversary with $\Cov=\V$ --- and ESA fails ($0\%$), concentrating frozen trust
mass $1.0$ onto it while the fail-safe cannot fire (it passes every audit). \emph{(c) Cardinal
elicitation is required:} re-running trust on preferences alone inverts the mechanism, $80\%\,[72,87]
\!\to\!0\%\,[0,0]$ (App.~\ref{app:proofs}, Rem.\ on modality).

\paragraph{Q6: Baselines on the real panel.} On the same real-grader panel (single 8/10 ratio), every
majority-trusting aggregator collapses: Mean $1\%$, Median/GAIL/Wass-DRO/RRM $0\%$, Dawid--Skene $1\%$;
KL-DRO reaches $38\%\,[29,48]$ --- its entropic risk exp-tilts toward low scores, discounting the
sycophant's lenient inflation, i.e.\ it exploits the same \emph{cardinal} channel as ESA but without a
reference --- while ESA recovers $80\%\,[72,87]$. A composition sweep (real graders) reproduces the
tipping point: aggregators hold until the sycophant fraction reaches $\approx 0.5$ and collapse
thereafter; ESA holds down to a single honest grader.

% =====================================================================
\section{Limitations}\label{sec:limits}
Source auditing cannot recover what the reference does not cover (Thm.~\ref{thm:lower}); with an
uncovered divergence region, ESA fails safe rather than correctly.

\paragraph{The defense is narrower than an audit suggests.} It holds against bias that is a
\emph{training artifact} --- dispositional sycophancy, whose leniency is verifier-visible and whose
cross-region coupling is therefore structural --- but not against an adversary \emph{strategic about
the audit boundary}. Against the latter no audit-based \emph{or} report-based rule can help
(Thm.~\ref{thm:lower} + Cor.~\ref{cor:internal}), and the failure is worse than inaction: because the
adversary is the \emph{best} grader on the audited region, ESA concentrates trust mass $1.0$ onto it
and the absolute-distrust fail-safe never fires (it passes every audit). In RLVR the coverage set is
\emph{public} (everyone knows math and code are checkable), so this adversary is the realistic
long-run case, not an exotic one --- and one cannot randomize or hide the audit region to defeat it.

\paragraph{Cardinal elicitation is a scope condition.} Every result assumes cardinal reports and a
cardinal reference; the dominance margin can go negative under purely ordinal (preference) elicitation,
and the mechanism then inverts (measured: trust mass $1.0$ on the sycophant, recovery $80\%\!\to\!0\%$).
This scopes ESA away from the modality of RLHF/DPO/WDPO/KLDPO and, conversely, argues for eliciting
absolute scores.

\paragraph{Other limits.} Aggregation flattens behavior: ESA changes \emph{whose} preferences dominate,
not the tendency of optimization to collapse onto a mode \citep[cf.][]{gao2023overopt}. Recovery is
bounded by the honest minority's own competence on the target region (empirically $\beta_\ast>0$: an
$80\%$ ceiling set by the honest graders' TruthfulQA accuracy). Transfer rests on bounded reliability
drift, which we measure but do not guarantee; verifiers are imperfect and gameable
\citep{skalse2022reward}. Formal results are bandit/contextual; MDP results are empirical.

% =====================================================================
\section{Conclusion}
Objective decoupling is structural, not incidental, and gold-standard consensus does not fix it once
bias is majority-held. Recovery by source auditing is possible exactly when the reference covers the
divergence region, and no cleverness on the reports alone removes that requirement. Verifiable
subtasks supply the reference for free where a ground truth exists and reduce the rest to a
measurable assumption. On real LLM graders this recovers $80\%$ of ground truth from a sycophantic
majority that defeats every consensus baseline --- but only for training-artifact bias, on cardinal
feedback, and only while the audit boundary is not itself the adversary's target.

\bibliographystyle{plainnat}
% \bibliography{yourbib}   % use your existing .bib + the entries added in
                           % ESA_corrections_and_upgrades.tex

\appendix
\section{Derivation of dominance}\label{app:dominance}\note{from corrections A5}
\section{Full proofs}\label{app:proofs}\note{Thms.~lower/upper/audit, Props.~failsafe/stackelberg/bloc/rlvr; from corrections file}
\section{Pseudocode}\label{app:algo}\note{ESA + ESA-UCB with active auditing, per-region trust, absolute distrust}
\end{document}
